Why is it that statistical tests for residuals are not widely used? An explanation using visual inference.

### 2023-01-24 paper_comparison/paper_comparison/paper_comparison.Rmd
fill in the last section of the result

  # Draw bootstrap CI for visual test
  geom_ribbon(data = ~get_boot_conf(.x,
                                    log_effect_size,
                                    reject,
                                    times = 500),
              aes(x = log_effect_size, 
                  ymin = lower, 
                  ymax = upper),
              fill = rev(rcartocolor::carto_pal(4, "Vivid"))[4],
              alpha = 0.2) +

@@

```{r polypower, fig.height = 3, fig.width = 5, fig.cap = "Comparison of power between different tests for non-linear patterns (uniform fitted values only). Main plot shows the power curves estimated using logistic regression, with dots indicating human evaluations of lineups. Surrounding band of the visual test shows the 95% confidence interval estimated using 500 bootstrap samples. Small row of plots shows typical residual plots corresponding to specific effect sizes, marked by dashed lines in main plot. Where would you draw the line of too much non-linearity in the residuals? For the RESET test this is around log effect size 1.5, but for the visual test it is around 3.5."}

@@

```{r heterpower, fig.height = 3, fig.width = 5, fig.cap = "Comparison of power between different tests for heteroskedasticity patterns (uniform fitted values only). Main plot shows the power curves, with dots indicating human evaluations of lineups. Surrounding band of the visual test shows the 95% confidence interval estimated using 500 bootstrap samples. Small row of plots shows typical residual plots corresponding to specific effect sizes, marked by dashed lines in main plot. Where would you draw the line of too much heteroskedasticity in the residuals? For the BP test this is around log effect size 1.5, but for the visual test it is around 3."}

@@

We also investigate the impact of the shape of heteroskedasticity on the power of different tests. The corresponding result is presented in Figure \ref{fig:heter-power-a}. In theory, the "Left-triangle" and the "Right-triangle" should have no difference from the point of view of a Breusch-Pagan test. According to Figure \ref{fig:heter-power-a}, this is indeed the case. And these two shapes should have the same visual test power, unless the orientation of the visual pattern is a non-negligible factor from human perspective. However, the same graph suggests that the "Right-triangle" shape is more difficult to be noticed by subjects than the "Left-triangle" shape.

@@

>>>>

  # Draw bootstrap CI for visual test
  geom_ribbon(data = ~get_boot_conf(.x,
                                    log_effect_size,
                                    reject,
                                    times = 1000),
              aes(x = log_effect_size, 
                  ymin = lower, 
                  ymax = upper),
              fill = rev(rcartocolor::carto_pal(4, "Vivid"))[4],
              alpha = 0.2) +

@@

```{r polypower, fig.height = 3, fig.width = 5, fig.cap = "Comparison of power between different tests for non-linear patterns (uniform fitted values only). Main plot shows the power curves estimated using logistic regression, with dots indicating human evaluations of lineups. Surrounding band of the visual test shows the 95 percent confidence interval estimated using 1000 bootstrap samples. Small row of plots shows typical residual plots corresponding to specific effect sizes, marked by dashed lines in main plot. Where would you draw the line of too much non-linearity in the residuals? For the RESET test this is around log effect size 1.5, but for the visual test it is around 3.5."}

@@

```{r heterpower, fig.height = 3, fig.width = 5, fig.cap = "Comparison of power between different tests for heteroskedasticity patterns (uniform fitted values only). Main plot shows the power curves, with dots indicating human evaluations of lineups. Surrounding band of the visual test shows the 95 percent confidence interval estimated using 500 bootstrap samples. Small row of plots shows typical residual plots corresponding to specific effect sizes, marked by dashed lines in main plot. Where would you draw the line of too much heteroskedasticity in the residuals? For the BP test this is around log effect size 1.5, but for the visual test it is around 3."}

@@

We also investigate the impact of the shape of heteroskedasticity on the power of different tests. The corresponding result is presented in Figure \ref{fig:heter-power-a}. In theory, the "Left-triangle" and the "Right-triangle" should have little difference from the point of view of a Breusch-Pagan test. According to Figure \ref{fig:heter-power-a}, this is indeed the case. And these two shapes should have the same visual test power, unless the orientation of the visual pattern is a non-negligible factor from human perspective. However, by checking the bootstrap confidence intervals, it can be observed from Figure \ref{fig:heter-power-a} that when $3 < log(E) < 4$, the "Left-triangle" power curve is significantly greater than the "Right-triangle" power curve. This phenomenon is particularly interesting. It implies the current residual plot design is not robust to the orientation of the visual pattern, which is worth to be explored in future research.

@@

# Conclusion

### 2023-01-24 paper_comparison/paper_comparison/paper_comparison.tex
fill in the last section of the result

\textbackslash begin\{figure\}

\{\centering \includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/polypower-1}

\}

\textbackslash caption\{Comparison of power between different tests for
non-linear patterns (uniform fitted values only). Main plot shows the
power curves estimated using logistic regression, with dots indicating
human evaluations of lineups. Surrounding band of the visual test shows
the 95\% confidence interval estimated using 500 bootstrap samples.
Small row of plots shows typical residual plots corresponding to
specific effect sizes, marked by dashed lines in main plot. Where would
you draw the line of too much non-linearity in the residuals? For the
RESET test this is around log effect size 1.5, but for the visual test
it is around 3.5.\}\label{fig:polypower} \textbackslash end\{figure\}

\textbackslash begin\{figure\}

\{\centering \includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/heterpower-1}

\}

\textbackslash caption\{Comparison of power between different tests for
heteroskedasticity patterns (uniform fitted values only). Main plot
shows the power curves, with dots indicating human evaluations of
lineups. Surrounding band of the visual test shows the 95\% confidence
interval estimated using 500 bootstrap samples. Small row of plots shows
typical residual plots corresponding to specific effect sizes, marked by
dashed lines in main plot. Where would you draw the line of too much
heteroskedasticity in the residuals? For the BP test this is around log
effect size 1.5, but for the visual test it is around
3.\}\label{fig:heterpower} \textbackslash end\{figure\}

@@

We also investigate the impact of the shape of heteroskedasticity on the
power of different tests. The corresponding result is presented in
Figure \ref{fig:heter-power-a}. In theory, the ``Left-triangle'' and the
``Right-triangle'' should have no difference from the point of view of a
Breusch-Pagan test. According to Figure \ref{fig:heter-power-a}, this is
indeed the case. And these two shapes should have the same visual test
power, unless the orientation of the visual pattern is a non-negligible
factor from human perspective. However, the same graph suggests that the
``Right-triangle'' shape is more difficult to be noticed by subjects
than the ``Left-triangle'' shape.

@@

>>>>

{\centering \includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/polypower-1}

\caption{Comparison of power between different tests for non-linear patterns (uniform fitted values only). Main plot shows the power curves estimated using logistic regression, with dots indicating human evaluations of lineups. Surrounding band of the visual test shows the 95 percent confidence interval estimated using 1000 bootstrap samples. Small row of plots shows typical residual plots corresponding to specific effect sizes, marked by dashed lines in main plot. Where would you draw the line of too much non-linearity in the residuals? For the RESET test this is around log effect size 1.5, but for the visual test it is around 3.5.}\label{fig:polypower}
\end{figure}

{\centering \includegraphics[width=1\linewidth]{paper_comparison_files/figure-latex/heterpower-1}

\caption{Comparison of power between different tests for heteroskedasticity patterns (uniform fitted values only). Main plot shows the power curves, with dots indicating human evaluations of lineups. Surrounding band of the visual test shows the 95 percent confidence interval estimated using 500 bootstrap samples. Small row of plots shows typical residual plots corresponding to specific effect sizes, marked by dashed lines in main plot. Where would you draw the line of too much heteroskedasticity in the residuals? For the BP test this is around log effect size 1.5, but for the visual test it is around 3.}\label{fig:heterpower}
\end{figure}

@@

We also investigate the impact of the shape of heteroskedasticity on the
power of different tests. The corresponding result is presented in
Figure \ref{fig:heter-power-a}. In theory, the ``Left-triangle'' and the
``Right-triangle'' should have little difference from the point of view
of a Breusch-Pagan test. According to Figure \ref{fig:heter-power-a},
this is indeed the case. And these two shapes should have the same
visual test power, unless the orientation of the visual pattern is a
non-negligible factor from human perspective. However, by checking the
bootstrap confidence intervals, it can be observed from Figure
\ref{fig:heter-power-a} that when \(3 < log(E) < 4\), the
``Left-triangle'' power curve is significantly greater than the
``Right-triangle'' power curve. This phenomenon is particularly
interesting. It implies the current residual plot design is not robust
to the orientation of the visual pattern, which is worth to be explored
in future research.

@@

\hypertarget{conclusion}{%
\section{Conclusion}\label{conclusion}}
